MLH1 Data Report

### 2019-09-10 doc/Main_Report.rmd
added files for lab meeting, work from home

@@

@@

>>>>

### Quality effect

there is a positive correlation with MLH1 number and quality
To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.

@@

## Explaination for the XX and XY differences

## Difference in proportion of chrm classes

generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
10 cells from each musc category.

```{r draw.rand.cells, echo=FALSE}
ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
```

```{r chrm_prop_data, echo=FALSE}

chrm.class.DF = read.csv("~./MLH1repo/ChrmClassProportions.csv", header = TRUE)
chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)

chrm.class.DF <- add_mouse(chrm.class.DF)
chrm.class.DF <- add_strain(chrm.class.DF)
chrm.class.DF <- add_sex(chrm.class.DF)
chrm.class.DF <- add_category(chrm.class.DF)
#chrm.class.DF <- add_euth_date(meta.data.file2)
#chrm.class.DF <- add_age(meta.data.file2)

table(chrm.class.DF$category)
str(chrm.class.DF)

#summarize by category
this_thing2  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                  tot_biv  = sum(biv.sum.check),

                 n0CO = sum(n0CO),
                  n1CO = sum(n1CO),
                  n2CO = sum(n2CO),
                  n3CO = sum(n3CO),
                  n4CO = sum(n4CO),

                 per0CO = sum(n0CO) / tot_biv,                 
                 per1CO= sum(n1CO)/ tot_biv,
                 per2CO= sum(n2CO)/ tot_biv,
                 per3CO= sum(n3CO)/ tot_biv,
                 per4CO= sum(n4CO)/ tot_biv,

                 mean.MLH1.foci = mean(MLH1.sum.check),
                 mean.MLH1.foci2 = mean(nMLH1.foci),
                 meanQ = mean(quality)

)
str(this_thing2)

#this first one is the 'first' MLH1 count
oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
summary(oopp2)
#none of the classes are significant

#mean is the new calculated-- mean 0CO is significant and intercept
oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
summary(oopp)
#not sure if these are inter-correlated
#I don't know why 0COs and the intercept are significant..
#0CO and 1CO are significant when only chrm classes are included (and intercept)

#no longer significant when used
#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
#after added more MSM data, tot_biv is not significant
#0CO is now more significant, and 2CO is sightly significant

#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias

#MSM needs more bialent observations

#use melt to put all propotions

#display / visualize / compare proportions

melt.table = read.csv("~./MLH1repo/ChrmClassProportions_melt.csv", header = TRUE)
melt.table$fileName <- (melt.table$Original.Name)

#melt.table <- add_mouse(melt.table)
#melt.table <- add_strain(melt.table)
#melt.table <- add_sex(melt.table)
#melt.table <- add_category(melt.table)

melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )

melt.table <- with(melt.table, melt.table[order(strain),])

other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp)

other_thing

chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check

chrm.class.DF$biv.sum.check

PWD and MSM have a higher proportion of 2CO chrms!

what about the variance across cells?

@@

## SC.skel measures

run analysis for whole SC skel measures

```{r RWscSkel, echo=FALSE}
#merge the sc.skel outputs with the master MLH1 measures
#match up random names with original names

#the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. I haven't manually confirmed that the default parameters are skeletonizing the images correctly.

#some of richard's analysis show negative relationship with SC length and MLH1 (his QTL word doc). Could this be sign
# of pwd like regulation?

setwd("~./")
SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_sep19.csv", header = TRUE, strip.white = TRUE)

SC.skel$rand_name <- as.character(SC.skel$rand_name)
original_DF$Random.Name <- as.character(original_DF$Random.Name)
str(SC.skel)
str(original_DF)

mlh1.skel.og <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
#for some reason this only worked after the chr and all = F

mlh1.skel.og <- mlh1.skel.og[ !grepl("X", mlh1.skel.og$Batch) , ]
mlh1.skel.og <- mlh1.skel.og[ !grepl("x", mlh1.skel.og$Batch) , ]
mlh1.skel.og <- mlh1.skel.og[!(is.na(mlh1.skel.og$quality) | mlh1.skel.og$quality==""), ]

mlh1.skel.og <- add_mouse(mlh1.skel.og)
mlh1.skel.og <- add_strain(mlh1.skel.og)
mlh1.skel.og <- add_sex(mlh1.skel.og)

mlh1.skel.og <- add_category(mlh1.skel.og)
mlh1.skel.og <- add_subsp(mlh1.skel.og)

#mlh1.skel.og <- add_euth_date(mlh1.skel.og)
#mlh1.skel.og <- add_age(mlh1.skel.og)

str(mlh1.skel.og)
mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)

#remove outliers
mlh1.skel <- mlh1.skel.og[mlh1.skel.og$skel_size < 25000,]
#first test file has ~1000 measures
#when I ran it, no sign of outliers (after removing all the X's cells)

looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=quality) ) + geom_jitter() + facet_wrap(sex~strain)

looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
#very possible that the dimesions and pixel density are skrewing with these values
# I am suspcious that there is such a wide range for the female data

looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))

#mouse level
mouse.level.skel <- ddply(.data=mlh1.skel.og, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(nMLH1.foci),
                 mean.skel = mean(skel_size)
)

mouse.level.skel <- add_strain(mouse.level.skel)
mouse.level.skel <- add_sex(mouse.level.skel)
mouse.level.skel <- add_category(mouse.level.skel)
mouse.level.skel <- add_subsp(mouse.level.skel)

looky.mouse <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Musc",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)

looky.mouseDom <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Dom",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)

#theme(legend.position="none")

#double check a way to visually inspect the python skel output
#
#repeat random numbers?
```

```{r test.BivData, echo=FALSE}
#bivdata
setwd("~./")
MSM_BivData.og = read.csv("~./MLH1repo/data/BivData/Curate_MSMmale_9.9.19.csv", header = TRUE, strip.white = TRUE )

MSM_BivData <- MSM_BivData.og[MSM_BivData.og$SC.pass ==1,]
MSM_BivData <- MSM_BivData[!(is.na(MSM_BivData$hand.foci.count) | MSM_BivData$hand.foci.count==""), ]

#MSM_BivDatam_CO2 <- MSM_BivData[MSM_BivData$hand.foci.count == 2,]
#MSM_BivDatam_CO2 <- MSM_BivDatam_CO2[!(is.na(MSM_BivDatam_CO2$fileName) | MSM_BivDatam_CO2$fileName==""), ]

#MSM_BivDatam_CO2$Foci1.PER <- MSM_BivDatam_CO2$Foci1 /MSM_BivDatam_CO2$chromosomeLength
#MSM_BivDatam_CO2$Foci2.PER <- MSM_BivDatam_CO2$Foci2 /MSM_BivDatam_CO2$chromosomeLength

MSM_BivData <- add_IFD(MSM_BivData)
#add IFD2 is f-ing up 
MSM_BivData <- addIFD.2(MSM_BivData)

#this MSM PER.IFD, is similar to the PWD data

#new function is converting all hand.foci counts to 3
#it might be the IFD functions

str(new_shiny)
str(MSM_BivData)

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
new_shiny <- add_SisCoTen(MSM_BivData)

new_shiny$rel_Siscoten <- new_shiny$SisCoTen / new_shiny$chromosomeLength

qq <- ggplot(data = new_shiny, aes(y=chromosomeLength, x=rel_Siscoten) ) + geom_jitter()

ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
  geom_boxplot(aes(alpha=.3))

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#KAZ and PWD data set is larger
PWD_KAZ_BivData.og = read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD.KAZ_BivData.csv", header = TRUE, strip.white = TRUE )

PWD_KAZ_BivData <- PWD_KAZ_BivData.og[PWD_KAZ_BivData.og$SC.pass == 1,]

PWD_KAZ_BivData <- add_IFD(PWD_KAZ_BivData)
PWD_KAZ_BivData <- addIFD.2(PWD_KAZ_BivData)
PWD_KAZ_BivData <- add_SisCoTen(PWD_KAZ_BivData)
PWD_KAZ_BivData$rel_Siscoten <- PWD_KAZ_BivData$SisCoTen / PWD_KAZ_BivData$chromosomeLength
PWD_KAZ_BivData <- add_mouse(PWD_KAZ_BivData)
PWD_KAZ_BivData <- add_strain(PWD_KAZ_BivData)

mouse_boxes <- ggplot(data = PWD_KAZ_BivData, aes(y=rel_Siscoten, x=mouse) ) + geom_jitter(aes( color=fileName))+
  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")

mouse_boxes2 <- ggplot(data = PWD_KAZ_BivData, aes(y=SisCoTen, x=mouse) ) + geom_jitter(aes( color=fileName))+
  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")

#mean relative and raw sisco-ten is higher in PWD compared to KAZ

biv_boxes <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count, fill=hand.foci.count) ) + geom_jitter()+geom_boxplot(aes(x= hand.foci.count,alpha=.3))+
theme(legend.position="none")+facet_wrap(~strain)

biv_boxes22 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=as.factor(hand.foci.count)) )+ geom_jitter()+
  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")+facet_wrap(~strain)

biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
theme(legend.position="none")+facet_wrap(~strain)

regress <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=chromosomeLength) ) + geom_jitter(aes( color=as.factor(hand.foci.count) ))+ facet_wrap(~strain)

regress.1 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
#, 1mar15_PWD_m1

regress.2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
#, 1mar15_PWD_m1

#

regress.1.KAZ <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$mouse == "21may18_KAZ_m2",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
 theme(legend.position="none")+ facet_wrap(~fileName)

#doesn't look like the sis-co-ten is conserved across bivs

```{r within.cell.corr, echo=FALSE}

#test telomere distance
#centromere distance
#SC length

#I want to test if SC lengths within cells cluster
#wouldn't ANOVA get at this..?

#cell level variance is significant predictor for SC length (telomere)

#PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",]
jjgg <- ddply(.data= PWD_KAZ_BivData,
                 .(fileName),
                 summarize, 
                 nbiv = length(unique(Obj.ID) ),
                  var.sc = var(chromosomeLength)
)

#reasons why this within cell correlation doesn't work well with bivalent data
  #different sets of chrms
#not enough resolution at cytology level

sis-coten metric:
I am currently missing a null model / test for this metric
(sis-co-ten from uniform placement of COs, and (some prob of draw for CO number))

there is a positive correlation with siscoten and chrm length

there could be something about kinetochore sites differing across specific chromosomes that balance the force
